Racial and Ethnic Differences in ETI: Outcomes

### 2025-08-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
nicer table formatting

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
    Years = -3:2,
    Race = unique(encounter$Race),
    age_group = "[18,30)",
    Insurance = "Private",
    prior_modulator = "Yes",
    isOnEnzymes = "Yes",
    pseudomonasaeruginosa = "No"
))
pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("FEV1 % Predicted") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt(caption = "Predicted Values at Various Timepoints") %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
# Zero-inflated models
# mod_p <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.poisson(), zi_fixed = ~1
# )
# mod_nb <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.negative.binomial(), zi_fixed = ~1
# )
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data,
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

@@

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

>>>>

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
    Years = -3:2,
    Race = unique(encounter$Race),
    age_group = "[18,30)",
    Insurance = "Private",
    prior_modulator = "Yes",
    isOnEnzymes = "Yes",
    pseudomonasaeruginosa = "No"
))
pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("FEV1 % Predicted") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(
        intercept = T,
        estimate_fun = ~ style_number(.x, digits = 1),
        pvalue_fun = ~ style_pvalue(.x, digits = 3)
    )
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt(caption = "Predicted Values at Various Timepoints") %>%
    fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(
        intercept = T,
        estimate_fun = ~ style_number(.x, digits = 1),
        pvalue_fun = ~ style_pvalue(.x, digits = 3)
    )
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(
        intercept = T,
        estimate_fun = ~ style_number(.x, digits = 1),
        pvalue_fun = ~ style_pvalue(.x, digits = 3)
    )
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(
        intercept = T,
        estimate_fun = ~ style_number(.x, digits = 1),
        pvalue_fun = ~ style_pvalue(.x, digits = 3)
    )
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 1)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
# Zero-inflated models
# mod_p <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.poisson(), zi_fixed = ~1
# )
# mod_nb <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.negative.binomial(), zi_fixed = ~1
# )
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data,
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```

@@

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 1)
```
